---
title: "Reproducible Research Project 1"
author: "chen"
date: "7/1/2020"
output:
  pdf_document: default
  html_document: default
---

## Loading and preprocessing the data

```{r, echo=TRUE, results = "hide"}
setwd("~/Coursera class/C5/project1")
act <- read.csv('./activity.csv')
head(act)
```

##### Libraries needed
```{r, results="hide"}
library(ggplot2)
library(mice)
```

## What is mean total number of steps taken per day?
##### aggregate to date level and then create the histogram
```{r}
day_steps <- tapply(act$steps, act$date, sum, na.rm=TRUE)
##day_steps <- aggregate(steps ~ date, data=act, FUN =sum, na.rm=TRUE)
hist(day_steps, breaks=20, main="Histogram of total number of steps taken each day")
```

The mean steps taken each day is **`r format(mean(day_steps))`**.

The median steps taken each day is **`r format(median(day_steps))`**.


## What is the average daily activity pattern?
1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
steps_by_int <- aggregate(steps ~ interval, data=act, FUN = mean, na.rm=TRUE)
plot(steps_by_int, type="l", xlab="interval", ylab="avg steps", main = "Activities at each interval across all days")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
steps_by_int$interval[which.max(steps_by_int$steps)]
```


## Imputing missing values

#####Use the mice package to impute missing using 'mean' method

```{r, echo=TRUE, results="hide"}
md.pattern(act)
act2 <- mice(act, m=5, maxit=50, meth='mean', seed=500)
completeact <- complete(act2,1)
```
Histogram of the total number of steps taken each day after missing values are imputed
```{r}
day_steps2 <- tapply(completeact$steps, completeact$date, sum, na.rm=TRUE)
hist(day_steps2, breaks=20, main="Histogram of total number of steps taken each day after imputing missing")
```

The mean steps taken each day after imputing missing is **`r format(mean(day_steps2))`**.

The median steps taken each day after imputing missing is **`r format(median(day_steps2))`**.

 
## Are there differences in activity patterns between weekdays and weekends?

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r, results="hide"}
completeact$daytype <- ifelse(as.POSIXlt(completeact$date)$wday %in% c(0,6), 'weekend', 'weekday')
table(completeact$daytype)
head(completeact)
```

```{r}
steps_by_int2 <- aggregate(steps ~ interval + daytype, data=completeact, FUN = mean, na.rm=TRUE)
g <- ggplot(steps_by_int2, aes(interval, steps))
g + geom_line() + facet_grid(daytype ~ .) + xlab("5- minte interval") + ylab("avg number of steps")
```

Yes, there is a difference in activity patterns between weekdays and weekends!
